### 2019-02-11 benchmarking/paper/draft0.Rmd
merged in benchmarking into v3b. changed 3.3v power pin so that design is compatible with teensy 3.1/2 and 3.5/6.

---
title: "Microstim: open source hardware for precise electrical stimulation"
author: ""
date: "February 11, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Microstim is built like so
  - Two independently-controllable output channels, each with both current and voltage output modes
  - Total cost: $131 USD.
  - Output range in current mode: -833uA to +833uA.
  - Pulse width in current mode: down to 50 us.
  - Output range in voltage mode: -12V to 12V.
  - Pulse width in current mode: down to 10 us.
  - In current mode, +-11V compliance voltage (+/-833uA output possible for resistances up to 13k, +-110uA possible for resistances up to 100k)
  - Channels are isolated 
  - Onboard ADC measurement of actual output current or voltage
  - runs on USB power
  - uses a Teensy-based microcontroller
  - FIGURE 1 (picture)

## Microstim is precise

  - amplitude accuracy
    - voltage and current output
  - pulse timing accuracy
  - channels are isolated
  - FIGURE 2 (benchmarks)

## Microstim is digital and highly configurable

## Microstim is better than everything else
  - TABLE 1 comparing stimulators?
    - STG3000 / 4000 from multichannel systems
    - WPI A310 Accupulser?
    - pulsepal

## Microstim limitations

## Potential applications

  - in vivo stimulation

## Supplemental materials
  - BOM
  - gerbers
  - schematic
  - link to bitbucket